Ecological performance of detecting data fabrication

### 2017-05-17 submission/manuscript.Rmd
Update study 1, add rmarkdown

---
title: "Ecological validity of detecting data fabrication in experimental studies."
author: "Chris HJ Hartgerink, Jelte M Wicherts, Marcel ALM van Assen"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: pdf_document
csl: ../bibliography/apa.csl
bibliography: ../bibliography/library.bib
---

```{r prepwork, echo = FALSE}
suppressPackageStartupMessages(if(!require(pROC)){install.packages('pROC')})
suppressPackageStartupMessages(library(pROC))
suppressPackageStartupMessages(if(!require(foreign)){install.packages('foreign')})
suppressPackageStartupMessages(library(foreign))

# Add ../ before running rmarkdown, or vice versa when not running there
x12 <- '../'
# x12 <- ''
```

<!---

# Abstract
--placeholder--

# Introduction

Data fabrication is malicious to scientific knowledge, hence it is imperative to be able to detect it. An estimated 2% of all researchers admit to having fabricated data at least once, and 14% have said to have witnessed someone fabricate data [@fanelli2009]. In stark contrast to plagiarism, which has a similar admittance rate but double the witness rate [@pupovac2015], hardly any systems are implemented during the editorial process to detect problematic data. Recently, one journal has explicitly stated they are considering such a system (Miller 2015), another is looking BMC SCIENCEMATTERS. Editors and reviewers hardly pay attention to potential problems with the data during the editor process either (Bornmann et al. 2008), which highlights the potential of automated detection systems.

Similar to how plagiarism scanners are implemented to detect plagiarism, statistical methods can be applied to detect potentially fabricated data. By extracting information from research papers, called content mining (Smith-Unna and Murray-Rust 2014), statistical analyses of the presented research results can be facilitated. Previous statistical content mining methods, extracting for example reported test results across 30,000 papers (Nuijten et al. 2015), indicate the feasibility of extracting statistical information from research papers in large quantities.

However helpful these technological advances are, the statistical methods applied to extracted data in order to detect potential data fabrication require thorough validation. More specifically, the statistical methods need to have a well-balanced trade-off between the false positive rate and the false negative rate. Considering the, presumably, low prevalence rate of data fabrication, false positive rates can have tremendous effects when applied at a large scale. For example, if we assume that 2% of the research papers is fabricated and we can detect it perfectly if it occurs, we would detect approximately 2.5 times more false positives than true positives. And this is if we assume the false positive rate is 5% and no higher.

How data is fabricated by researchers is not well-known and simulating the properties of statistical methods to detect data fabrication is therefore invalid. When the process is unknown, it is hardly feasible to properly approximate reality and draw valid conclusions. Moreover, it seems reasonable to state that there are many different ways to fabricate data and that these different data fabrication ‘styles’ can be combined. Even if certain fabrication processes were known, simulating ecologically valid processes would still be quite difficult.

Nonetheless, statistical methods to detect data fabrication have been developed based on theoretical principles, which help detect data fabrication in experiments. These methods are based on the tenet that humans are bad at fabricating results that originate from probabilistic processes. Outcome variables that are inspected in experiments are based on a sample which results in observations that are the result of a random variable, hence subject to probabilistic (i.e., stochastic) processes. The inadequacy of humans in generating such sequences is the consequence of mental heuristics. For example, most humans succumb to the gambler’s fallacy, estimating the probability of heads as more likely when we have thrown tails several times (Tversky and Kahneman 1974). Other fallacious heuristics include the local-representativeness heuristic (), .

Statistical methods to detect data fabrication have been successfully applied in ad hoc investigations, but their value in screening processes remains opaque. For example, Simonsohn’s variance of variances method (Simonsohn 2013) has been applied to the Smeesters and Sanna cases, but only after suspicions had already arisen. The value of applying such a method as a screening measure is hard to assess regardless of its previous successes in ad hoc investigations.

In this paper, we experimentally test the validity of a set of statistical methods to detect data fabrication, based on summary results (Study 1) or raw data (Study 2), and assess the optimal balance between decision criterion and detection accuracy. We apply the statistical methods to detect data fabrication to both (assumed) genuine data from the Many Labs experiments and fabricated datasets, which we collected in an experimental setting where researchers were invited to fabricate data according to a controlled set of hypotheses. Both experiments were approved by the Tilburg Ethics committee (EC-2015.50).
--->
# Study 1

```{r, echo = FALSE}
# Dynamically generate path names
ml_file <- sprintf('%sdata/study_01/ml_summary_stats.csv', x12)
qualtrics_file <- sprintf('%sdata/study_01/qualtrics_processed.csv', x12)
dat_summary_file <- sprintf('%sdata/study_01/study1_res.csv', x12)
file <- sprintf('%sdata/study_01/raw_summary_results_fabrication_qualtrics.csv', x12)
res_file <- sprintf('%sdata/study_01/qualtrics_processed.csv', x12)
ml_dat_file <- sprintf('%sdata/study_01/anchoring_ml/chjh ml1_anchoring cleaned.sav', x12)
summary_stat_file <- sprintf('%sdata/study_01/ml_summary_stats.csv', x12)
pdf_file <- sprintf('%sarchive/gender_interaction.pdf', x12)

@@

# Compute the summary statistics for Many Labs
source(sprintf('%sfunctions/compute_summary_anch_ml.R', x12))
# Process the collected, fabricated data
source(sprintf('%sfunctions/process_qualtrics_anch_01.R', x12))
# Concatenate, analyze, and write out all ML and qualtrics data
# Uncomment this line to rerun
# set.seed(123);source(sprintf('%sfunctions/concatenate_analyze_01.R', x12))
```

We tested the performance of statistical methods to detect data fabrication based on summary results with genuine and fabricated summary results of four anchoring studies [@tversky1974;@jacowitz1995]. The anchoring effect is a well-known psychological heuristic that uses the information in the question as the starting point for the answer, which is then adjusted to yield a final estimate of a quantity. For example 'Is the percentage of African countries in the United Nations more or less than [10\% or 65\%]?'. These questions yield mean responses of 25\% and 45\%, respectively [@tversky1974], despite essentially posing the same factual question. A considerable amount of genuine datasets on this heuristic are freely available and we collected fabricated datasets within this study.

@@

The four anchoring studies for which results were collected were (i) distance from San Francisco to New York, (ii) population of Chicago, (iii) height of the Mount Everest, and (iv) the number of babies born per day in the United States. Each of the four studies provided summary results for a 2 (low/high anchoring) × 2 (male/female) factorial design. Throughout this study, the unit of analysis is a set of summary statistics (i.e., means, standard deviations, and test results) for the four anchoring studies from one respondent. Respondent is defined as researcher/lab where the four anchoring studies' summary statistics originate from. All materials, data, and analyses scripts are freely available on the OSF ([osf.io/b24pq](osf.io/b24pq)) and were preregistered ([osf.io/ejf5x](osf.io/ejf5x); deviations are explicated in this report).

### Data collection.

We downloaded thirty-six genuine datasets from the publicly available Many Labs (ML) project [@klein2014, [osf.io/pqf9r](osf.io/pqf9r)]. The ML project replicated several effects across thirty-six locations, including the anchoring effect in the four studies mentioned previously. Considering the size of the ML project, the transparency of research results, and minimal individual gain for fraud, we assumed these data to be genuine. For each of the thirty-six locations, sample sizes, means, and standard deviations (four each) were computed, for each of the four conditions in the four anchoring studies across the thirty-six locations (i.e., $3\times4\times4\times36$). We computed these summary statistics from the raw ML data, which were cleaned using the original analysis scripts from the ML project.

Using quotum sampling, we collected thirty-six fabricated datasets of summary results for all four anchoring studies. The sampling frame consisted of 2,038 psychology researchers who published a peer-reviewed paper in 2015, as indexed in the Web of Science (WoS) with the filter set to the U.S. We sampled psychology researchers to improve familiarity with the anchoring effect [@jacowitz1995;@tversky1974], for which summary results were fabricated. We filtered for U.S. researchers to ensure familiarity with the imperial measurement system, which is the scale of some of the anchoring studies (note: we found out several non-U.S. researchers were included because this filter also retained papers with co-authors from the U.S.). WoS was searched on October 13, 2015. In total, 2,038 unique corresponding e-mails were extracted from 2,014 papers (due to multiple corresponding authors).

A random sample of 1,000 researchers were digitally approached to participate in this study on April 25, 2016 (invitation: [osf.io/s4w8r](osf.io/s4w8r)). The study took place via Qualtrics with anonimization procedures in place (e.g., no IP-addresses saved). We informed the participating researchers that the study would require them to fabricate data and explicitly mentioned that we would investigate these data with statistical methods to detect data fabrication. We also clarified to the  respondents that they could stop at any time without providing a reason. If they wanted, respondents received a $30 Amazon gift card as compensation for their participation for which they had to provide their email address, after which they could win an additional $50 Amazon gift card if they were one of three top fabricators. The provided email addresses were unlinked from individual responses upon sending the bonus gift cards. The text of the Qualtrics survey is available at [osf.io/w984b](osf.io/w984b).

Each respondent was instructed to fabricate 32 summary statistics (4 studies $\times$ 2 conditions $\times$ 2 sexes $\times$ 2 statistics [mean and sd]) that fulfilled three hypotheses. We instructed respondents to fabricate results for the hypotheses (i) main effect of condition, (ii) no effect of sex, and (iii) no interaction effect between condition and sex. Respondents did not need to fabricate sample sizes, which were set to 25 per cell a priori. The fabricated summary statistics and their accompanying test results for these three hypotheses serve as the data to examine the properties of tools to detect data fabrication.

We provided respondents with a template spreadsheet to fill out the fabricated data, in order to standardize the fabrication process without restraining the participant in how they choose to fabricate data. Figure `r fig <- 1; fig` depicts an example of this spreadsheet (original:  [osf.io/w6v4u](osf.io/w6v4u)). We requested the respondents to fill in the yellow cells with fabricated data, which includes means and the standard deviations for four conditions. Using these values, statistical tests are computed and shown in the "Current result" column instantaneously. When these results confirm the hypotheses, a checkmark appears as depicted in the green cells. We required respondents to copy-paste the yellow cells into Qualtrics, to provide a standardized response format that could be automatically processed in the analyses.

![spreadsheet](../figures/spreadsheet.png)

**Figure `r fig`.** Example of a filled in template spreadsheet used in the fabrication process. Respondents fabricated data in the yellow cells, which were used to compute the results of the hypothesis tests. If the fabricated data confirm the hypotheses, a checkmark appeared in a green cell.

Upon completing the fabrication of the data, respondents were debriefed. Several questions were asked about their statistical knowledge and approach to data fabrication. They were also reminded that data fabrication is widely condemned by professional organizations, institutions, and funding agencies alike.

We rewarded participation with a $30 Amazon gift card and the fabricated results that were most difficult to detect received a bonus $50 Amazon gift card. If the participant wanted to receive a compensation and contend for the bonus $50, he/she had to enter an email to receive the reward. These email addresses were unlinked from individual responses upon sending the gift cards. Quotum sampling was applied to sample as many responses as possible for the available 36 rewards (i.e., not all respondents might request the gift card and count towards the quotum; one participant did not request a reward).

### Data analysis.

To detect data fabrication in a set of summary results, we first test the variance of fabricated standard deviations [SDs; @simonsohn2013] across the four anchoring studies. This method tests whether the observed SDs contain a reasonable amount of variation, as expected based on stochastic sampling processes. For example, if four independent samples all yield the variance of 2.22, this could be considered excessively consistent when the probability that this amount of consistency (or more), given genuine samples, is less than 1 out of 1000 (this is simply an example). To compute this test, we first standardize the SDs for each of the four studies by computing

$\tilde{s}^2_j=\frac{s^2_j}{MS_w}=\frac{s^2_j}{\frac{\sum(N_j-1)s^2_j}{\sum(N_j-1)}}$
<!--- \limits^k_{j=1} --->

where $\tilde{s}_j$ denotes the standardized SD in group *j*. Note that $MS_w$ is the simple arithmetic mean when sample sizes are equal for all cells. We test several different measures to detect data fabrication that utilize these standardized SDs (i.e., $\tilde{s}_j$), including the max-min distance (denoted $\tilde{\sigma}_{max-min}$) and the variance of the standardized SDs [i.e.,  $\tilde{\sigma}_{sd}$; @simonsohn2013]. We compare the observed value for each $\tilde{\sigma}$ measure with the expected distribution of outcomes when the data would be the result of random sampling processes. To this end, we simulate the expected distribution of standardized SDs and compute the expected distribution of each $\tilde{\sigma}$. This expected distribution is used to determine the *p*-value of the observed $\tilde{\sigma}$ value. We simulate the standardized SD for each of the *j* groups as

$\tilde{s}^2_j\sim\Bigg[\frac{\chi^2_{N_j-1}}{N_j-1}\Bigg]/MS_w$

These simulated values are used to compute the expected distribution of all $\tilde{\sigma}$ measures.

>>>>

---
title: "Ecological performance of detecting data fabrication"
author: "Chris HJ Hartgerink, Jan G Voelkel, Jelte M Wicherts, Marcel ALM van Assen"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  word_document: default
  html_document:
    toc: yes
csl: ../bibliography/apa.csl
bibliography: ../bibliography/library2.bib
---

```{r preparation, echo = FALSE}
suppressPackageStartupMessages(if(!require(pROC)){install.packages('pROC', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(pROC))
suppressPackageStartupMessages(if(!require(foreign)){install.packages('foreign', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(foreign))
suppressPackageStartupMessages(if(!require(latex2exp)){install.packages('latex2exp', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(latex2exp))
suppressPackageStartupMessages(if(!require(ggplot2)){install.packages('ggplot2', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(if(!require(gridExtra)){install.packages('gridExtra', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(if(!require(xtable)){install.packages('xtable', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(if(!require(httr)){install.packages('httr', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(httr))
suppressPackageStartupMessages(if (!require(car)){install.packages('car', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(if (!require(effects)){install.packages('effects', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(effects))
suppressPackageStartupMessages(if (!require(lsr)){install.packages('lsr', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(lsr))
suppressPackageStartupMessages(if (!require(data.table)){install.packages('data.table', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(if (!require(dplyr)){install.packages('dplyr', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(if (!require(reshape2)){install.packages('reshape2', repos = 'http://cran.us.r-project.org')})
suppressPackageStartupMessages(library(reshape2))

# Add ../ before running rmarkdown, or vice versa when not running there
x12 <- '../'
# x12 <- ''

source(sprintf('%sfunctions/digit_analysis.R', x12))
source(sprintf('%sfunctions/std_var.R', x12))

ml_file <- sprintf('%sdata/study_01/ml_summary_stats.csv', x12)
qualtrics_file <- sprintf('%sdata/study_01/qualtrics_processed.csv', x12)
dat_summary_file <- sprintf('%sdata/study_01/study1_res.csv', x12)
file <- sprintf('%sdata/study_01/raw_summary_results_fabrication_qualtrics.csv', x12)
res_file <- sprintf('%sdata/study_01/qualtrics_processed.csv', x12)
ml_dat_file <- sprintf('%sdata/study_01/anchoring_ml/chjh ml1_anchoring cleaned.sav', x12)
summary_stat_file <- sprintf('%sdata/study_01/ml_summary_stats.csv', x12)
ml3_dat_file <- sprintf('%sdata/study_02/study_02-ml3_stroop/StroopCleanSet.csv', x12)
pdf_file <- sprintf('%sarchive/gender_interaction.pdf', x12)

@@

tabnr <- 1
fignr <- 1
```

Any field of empirical inquiry is faced with cases of scientific misconduct at some point, either in the form of fabrication, falsification or plagiarism (FFP). Psychology was faced with Stapel; medical sciences were faced with Poldermans and Macchiarini; life sciences were faced with Voignet. These are just a few examples of misconduct cases in the last decade. Overall, an estimated 2\% of all scholars have admitted to falsifying or fabricating research results at least once [@10.1371/journal.pone.0005738] and this is likely to be an underestimate due to socially desirable responses. The detection rate is likely to be even lower; for example, only around a dozen cases are discovered in the United States and the Netherlands, despite covering several hundreds of thousands of researchers. At best, this amounts to a detection rate far below 1\% of those 2\% who admit to fabricating data --- the tip of a seemingly much larger iceberg.

In order to stifle attempts at data fabrication, improved detection of fabricated data is considered to deter such harmful attempts. 
Although deterrence theory dates back to the middle of the 17th century [@leviathan], its implementation has not occurred across the different forms of scientific misconduct equilaterally. Basically, deterrence theory stipulates that with increased risk of detection, the utility of scientific misconduct (for this context) will decrease and therefore fewer people will engage in such behaviors. This principle of deterrence has been implemented with plagiarism scanners, a development that already started a long time ago [e.g., @10.1109/13.28038]. However, increased deterrence of fabrication and falsification by improved detection mechanisms has not been as widely implemented.

In the last decade, detecting image manipulation has become one of the few forms of detecting scientific misconduct (alongside plagiarism).
The Journal of Cell Biology scans each submitted image for potential manipulation [@The_Journal_of_Cell_Biology2015-vh], which greatly increases the risk of detecting (blatant) image manipulation. More recently, algorithms have been developed to automate the scanning of images for (subtle) manipulations [@10.1007/s11948-016-9841-7]. These developments in detecting image manipulation have increased detection risk during the pre-publication and post-publication phase by improving detection mechanisms and increasing the understanding of how images might be manipulated. Moreover, their application also helps researchers systematically evaluate research articles to estimate the extent of the problem of image manipulation [4\% of all papers are estimated to contain manipulated images, @10.1128/mBio.00809-16].

Statistical methods can provide one way to improve detection of data fabrication in empirical research. Humans are notoriously bad at understanding and estimating probabilities [e.g., @10.1126/science.185.4157.1124;@10.1037/h0031322], which could manifest itself in the fundamentally probabilistic data they try to fabricate. 
That researchers do not understand probabilistic processes also presents itself in false interpretation of genuine research data [@10.3758/BF03213921;@10.1007/s11336-015-9446-0;@10.7326/0003-4819-130-12-199906150-00008]. When data are fabricated, probabilistic principles are easily violated if forgotten at the univariate level, bivariate level, trivatiate level, or beyond [@Haldane1948-nm]. Based on this idea, statistical methods that investigate whether the reported data are feasible under the theoretically probabilistic processes can be used to detect potential data fabrication.

The application of such statistical methods to detect data fabrication has occurred in several cases in recent years and has potential for future application beyond a case-basis. 
For example, problems in papers by Fuji were highlighted with statistical methods [@10.1111/j.1365-2044.2012.07128.x;@10.1111/anae.13126], resulting in 183 retractions [@oransky2015]. In this case, baseline measures across randomized groups were examined for too little variation. Random assignment should introduce a certain degree of random error that might be missed by a human fabricator, misestimating the probabilistic process that generates such error. Another two cases are those of Sanna and Smeesters, where fabricated data were also detected with statistics [@10.1177/0956797613480366]. These cases inspected the variance of variances (i.e., the second, or meta level variance). Once again, too little variation was what revealed problems in these data [other examples include @lacour-complaint;@foerster-complaint]. These methods, although developed in a case-setting, need not be limited to cases. The application of such methods can be (semi-)automated if data are available in a machine-readable format that one of the statistical methods can be applied to. An example of such a potential case for mass application of using statistics to detect (potential) data fabrication is in the ClinicalTrials.gov database, where baseline measures across randomized groups are readily available for download and subsequent analysis [@10.3897/rio.1.e7462].

Nonetheless, considering the potential harm of applying statistical methods to flag potentially problematic results, it needs to be sorted out whether such methods have diagnosticity that actually makes it responsible to apply them. 
We hardly know how researchers might go about fabricating data. Cases such as Fuji, Smeesters, and Sanna provide some insights, but are highly pre-selected (i.e., those who got caught/confessed) and as such, systematically biased. Relatively extensive descriptions in rare and partial autobiogrophical accounts provide little insight into the actual data fabrication process, except for the setting where it might take place [e.g., late at night when no one is around; @stapel_book]. However, trustworthiness of these accounts can be called into question. Additionally, the performance of methods to detect data fabrication is highly dependent on the unknown prevalence of data fabrication and the power to actually to detect data fabrication. Given that we do not know how researchers might fabricate data, the diagnosticity of these methods cannot realistically be simulated.

The effectiveness of detection mechanisms and their consequences, hence their expected deterrence, is exacerbated in the digital age by the increased usage of public online discussion platforms such as [PubPeer (https://pubpeer.com)](https://pubpeer.com). 
PubPeer serves as an "online journal club" where anyone can discuss articles. Authors are notified when someone leaves a response, providing them with the possibility to respond. Such a platform allows for public discussion of the paper, including discussion of reanalyses, methodology, availability of materials, etc. When detection mechanisms are freely available to use, they can lead to (a surge of) comments when applied by users. Recently, in 2016, one user (the main author of this paper) used `statcheck` software to report potential statistical reporting inconsistencies for 50,000 psychology articles, which led to a large discussion about (automated) online comments [@10.1038/540151a]. The impact of such new possibilities is not to be underestimated, although its potential to constructively contribute to the scientific discussion should also not be [@10.1038/540007b]. Nonetheless, the `statcheck` software was well validated prior to this application [@10.3758/s13428-015-0664-2] and the same principle applies to the application of statistical methods to detect (potential) data fabrication.

# Theoretical framework

Throughout this paper, we inspect statistical methods to detect data fabrication that can be applied to (1) summary results or (2) raw data. 
Even though structure and contents of data can look different depending on the structure of a study and the measures, there are certain common characteristics of empirical results and the underlying raw data that can be inspected. For example, summary results frequently include means, standard deviations, test-statistics, and $p$-values. Raw data frequently contain at least some variables measured at a interval- or ratio scale [@10.1126/science.103.2684.677]. Such common characteristics allow for the development of generic statistical methods that can be applied across a varied set of results. We review the theoretical framework of the specific methods we apply throughout this paper, but these are not exhaustive of all methods available to test for potential problems in empirical data [see also @buyse1999;@10.7287/peerj.preprints.2400v1;@10.7287/peerj.preprints.2064v1;@sprite-heathers].

## Detecting data fabrication in summary results

### P-value analysis

The distribution of one $p$-value depends on various parameters, but there are specific theoretical boundary conditions to this distribution [@fisher1925]. Based on the population effect size, the precision of the estimate, and the observed effect size, the $p$-value distribution can be uniform (when then null hypothesis is true) or right-skewed (when the alternative hypothesis is true). By extension, any other distribution, after taking into account sampling error, is theoretically suspect. For example, a distribution might show a bump when authors $p$-hack [e.g., Figure 1 in @10.7717/peerj.1935], but fabricators might create other non-uniform or non-right-skewed distributions, failing to take into account the theoretical boundary conditions.

In order to test whether observed $p$-values transgress the boundary condition, we proposed an adaptation of Fisher's method [@10.1186/s41073-016-0012-9]. This adaptation is a simple reversal of the Fisher method [@fisher1925], which was originally introduced as a simple meta-analytic test to indicate presence of an effect; this test is computed as 
$$
\chi^2_{2k}=-2\sum\limits^k_{i=1}\ln(p_i)
$$
where it tests for right-skew (i.e., more smaller $p$-values than larger $p$-values) across the $k$ number of $p$-values. Reversing the Fisher method results in 
$$
\chi^2_{2k}=-2\sum\limits^k_{i=1}\ln(1-\frac{p_i-t}{1-t})
$$
where it now tests for left-skew (i.e., more larger $p$-values than smaller $p$-values) across the $k$ number of $p$-values that falls above the threshold $t$. This threshold is added and can be set to anything. Given the frequent misinterpretation of $p$-values [e.g., as the probability of an effect, @10.1053/j.seminhematol.2008.04.003;@10.1136/bmj.311.7003.485], researchers who fabricate nonsignificant data might forget to fabricate a uniform $p$-value distribution and generate only large $p$-values when fabricating nonsignificant results. In that case, it makes  sense to set $t=.05$ or $t=.10$, in order to include only nonsignificant test results.

Despite this test being useful for detecting data anomalies in nonsignificant $p$-values, two clear exceptions should be taken into account: (1) model-fit tests and (2) wrongly specified one-tailed tests. When model-fit tests are used, these can quickly result in high $p$-values because of model saturation (i.e., where the null hypothesis is good fit and the alternative is bad fit). For properly specified one-tailed tests, the $p$-value distribution is right-skewed. When wrongly specified, this distribution is reversed and becomes left-skew. These two exceptional cases should not be forgotten during the application of this test, but are irrelevant for the rest of this paper due to our study design (see methods of both studies).

### Variance analysis

Variance- or standard deviation estimates are typically reported to indicate spread, but just like the mean there should be sampling error in this estimate (i.e., the variance of the variances) proportionate to the sample size. A variance estimate follows a $\chi^2$-distribution, which is dependent on the sample size [p. 445; @hogg-tanis]; that is
$$
z^2_j\sim\left(\frac{\chi^2_{N_j-1}}{N_j-1}\right)/MS_w
$$
where $N_j$ is the sample size of the $j$th group and $MS_w$ is the normalizing constant resulting in a standardized variance $z^2_j$. The normalizing constant $MS_w$ is computed as 
$$
MS_w=\frac{\sum\limits^k_{j=1}(N_j-1)s^2_j}{\sum\limits^k_{j=1}(N_j-1)}
$$
where $s^2_j$ is the variance in the $j$th group. If simulating (see next paragraph), the simulated variances are used to compute $MS_w$.

Assuming that the observed variances are from the same population distribution, the expected spread of reported standardized variances can be simulated. By repeatedly drawing (standardized) variance estimates for $j$ groups and then computing their spread allows for approximation of the sampling distribution. Spread can be operationalized in various ways, such as the standard deviation of the variances (denoted in this paper as $SD_z$) or as the range of the variances (denoted as $max-min_z$). If there is reason to believe that the observed variances come from heterogeneous populations, subgrouping (i.e., one variance analysis per subgroup) is necessary to prevent false results.

Subsequently, the observed sampling variance (or range) can be compared to the expected sampling distribution. Too consistent results would indicate potential anomalies in the reported data [@10.1177/0956797613480366]. For example, if four independent samples all yield the variance $2.22$, this could be considered excessively consistent when the probability that this amount of consistency (or more) is less than 1 out of 1000 in truly random samples.

### Effect size analysis

A fabricator might present unrealistically large effects, forgetting the exponential relation between the effect size measure and the implied correlation. From our own experience, and anecdotal evidence elsewhere [@10.1016/0197-24569190037-M], large effects have previously raised initial suspicions. The size of a large effect can quickly become unrealistic; $d=1.2$ already implies a correlation of `r round(sqrt((1.2^2/(1.2^2+4))), 2)` between the dependent- and independent measures. More generally, the relation between the implied correlation and the effect size tends to be exponential, as Figure `r fignr` depicts.

```{r d2, fig.cap="_Figure XX._ The relation between Cohen's d effect size and its direct transformation, the correlation r (or vice versa). If effects d>2 are fabricated, the implied correlation between the dependent- and independent measure(s) is approximately .6.", out.width='100%', echo=FALSE}
d2 <- seq(0, 10, .01)

plot(sqrt(d2^2/(d2^2+4)), x = d2,
     type='l', xlab = "Cohen's d", ylab = "Correlation r")

fignr <- fignr + 1
```

Taking the observed effect size and transforming it into a correlation, allows for an easy way to assess how extreme the presented result is. One minus the observed correlation can be used as a measure for extreme effects (i.e., $1-r$); as a heuristic, it can be regarded as a $p$-value. That is, this measure too ranges from zero to one and the more extreme the effect size, the smaller the value. This method specifically looks at situations where fabricators would want to fabricate the existence of an effect (not the absence of one).

## Detecting data fabrication in raw data

### Digit analysis

### Variance analysis

### Multivariate associations

# Study 1

We tested the performance of statistical methods to detect data fabrication in summary results with genuine- and fabricated summary results from four anchoring studies [@10.1126/science.185.4157.1124;@10.1037/e722982011-058]. The anchoring effect is a well-known psychological heuristic that uses the information in the question as the starting point for the answer, which is then adjusted to yield a final estimate of a quantity. For example 'Is the percentage of African countries in the United Nations more or less than [10\% or 65\%]?'. These questions yield mean responses of 25\% and 45\%, respectively [@10.1126/science.185.4157.1124], despite essentially posing the same factual question. A considerable amount of genuine datasets on this heuristic are freely available and we collected fabricated datasets within this study. This study was approved by the Tilburg Ethical Review Board (EC-2015.50).

@@

```{r prep study 1, echo=FALSE}
# Study 1
# Compute the summary statistics for Many Labs
suppressWarnings(suppressMessages(source(sprintf('%sfunctions/compute_summary_anch_ml.R', x12))))
# Process the collected, fabricated data
suppressMessages(source(sprintf('%sfunctions/process_qualtrics_anch_01.R', x12)))
# Concatenate, analyze, and write out all ML and qualtrics data
# Uncomment this line to rerun, but take care: might take a while
if(!file.exists(sprintf('%sdata/study_01/study1_res.csv', x12)))
{
  set.seed(123);suppressMessages(source(sprintf('%sfunctions/concatenate_analyze_01.R', x12)))  
}
```

We collected summary results for four anchoring studies: (i) distance from San Francisco to New York, (ii) population of Chicago, (iii) height of the Mount Everest, and (iv) the number of babies born per day in the United States [@10.1037/e722982011-058]. Each of the four studies provided us with summary results for a 2 (low/high anchoring) $\times$ 2 (male/female) factorial design. Throughout this study, the unit of analysis is a set of summary statistics (i.e., means, standard deviations, and test results) for the four anchoring studies from one respondent. For current purposes, a respondent is defined as researcher/lab where the four anchoring studies' summary statistics originate from. All materials, data, and analyses scripts are freely available on the OSF ([https://osf.io/b24pq](https://osf.io/b24pq)) and a preregistration is available at [https://osf.io/ejf5x](https://osf.io/ejf5x) (deviations are explicated in this report).

### Data collection

We downloaded thirty-six genuine datasets from the publicly available Many Labs (ML) project [[https://osf.io/pqf9r](https://osf.io/pqf9r); @10.1027/1864-9335/a000178]. The ML project replicated several effects across thirty-six locations, including the anchoring effect in the four studies mentioned previously. Considering the size of the ML project, the transparency of research results, and minimal individual gain for fabricating data, we assumed these data to be genuine. For each of the thirty-six locations we computed sample sizes, means, and standard deviations for each of the four conditions in the four anchoring studies (i.e., $3\times4\times4$) for each of the thirty-six locations. We computed these summary statistics from the raw ML data, which were cleaned using the original analysis scripts from the ML project.

Using quotum sampling, we collected thirty-six fabricated datasets of summary results for the same four anchoring studies. Quotum sampling was used to sample as many responses as possible for the available 36 rewards (i.e., not all respondents might request the gift card and count towards the quotum; one participant did not request a reward). The sampling frame consisted of 2,038 psychology researchers who published a peer-reviewed paper in 2015, as indexed in Web of Science (WoS) with the filter set to the U.S. We sampled psychology researchers to improve familiarity with the anchoring effect [@10.1126/science.185.4157.1124;@10.1037/e722982011-058], for which summary results were fabricated. We filtered for U.S. researchers to ensure familiarity with the imperial measurement system, which is the scale of some of the anchoring studies (note: we found out several non-U.S. researchers were included because the WoS filter also retained papers with co-authors from the U.S.). WoS was searched on October 13, 2015. In total, 2,038 unique corresponding e-mails were extracted from 2,014 papers (due to multiple corresponding authors).

We invited a random sample of 1,000 researchers via e-mail to participate in this study on April 25, 2016 (invitation: [https://osf.io/s4w8r](https://osf.io/s4w8r)). The study took place via Qualtrics with anonimization procedures in place (e.g., no IP-addresses saved). We informed the participating researchers that the study would require them to fabricate data and explicitly mentioned that we would investigate these data with statistical methods to detect data fabrication. We also clarified to the  respondents that they could stop at any time without providing a reason. If they wanted, respondents received a $30 Amazon gift card as compensation for their participation if they were willing to enter their email address. They could win an additional $50 Amazon gift card if they were one of three top fabricators. The provided e-mail addresses were unlinked from individual responses upon sending the bonus gift cards. The full text of the Qualtrics survey is available at [https://osf.io/w984b](https://osf.io/w984b).

Each respondent was instructed to fabricate 32 summary statistics (4 studies $\times$ 2 conditions $\times$ 2 sexes $\times$ 2 statistics [mean and sd]) that fulfilled three hypotheses. We instructed respondents to fabricate results for the following hypotheses: there is (i) a main effect of condition, (ii) no effect of sex, and (iii) no interaction effect between condition and sex. We fixed the sample sizes to 25 per cell; respondents did not need to fabricate sample sizes. The fabricated summary statistics and their accompanying test results for these three hypotheses serve as the data to examine the properties of statistical tools to detect data fabrication.

We provided respondents with a template spreadsheet to fill out the fabricated data, in order to standardize the fabrication process without restraining the participant in how they chose to fabricate data. Figure `r fignr` depicts an example of this spreadsheet (original:  [https://osf.io/w6v4u](https://osf.io/w6v4u)). We requested respondents to fill in the yellow cells with fabricated data, which includes means and the standard deviations for four conditions. Using these values, statistical tests are computed and shown in the "Current result" column instantaneously. If these results confirmed the hypotheses, a checkmark appeared as depicted in Figure `r fignr`. We required respondents to copy-paste the yellow cells into Qualtrics, to provide a standardized response format that could be automatically processed in the analyses.

```{r spreadsheet-study1, fig.cap="_Figure XX._ Example of a filled in template spreadsheet used in the fabrication process of Study 1. Respondents fabricated data in the yellow cells, which were used to compute the results of the hypothesis tests. If the fabricated data confirm the hypotheses, a checkmark appeared in a green cell (one of four template spreadsheets available at [https://osf.io/w6v4u/](https://osf.io/w6v4u/)).", out.width='100%', echo=FALSE}
knitr::include_graphics('../figures/spreadsheet.png')

fignr <- fignr + 1
```

Upon completing the fabrication of the data, respondents were debriefed. Respondents answered several questions about their statistical knowledge and approach to data fabrication and finally we reminded them that data fabrication is widely condemned by professional organizations, institutions, and funding agencies alike. We rewarded participation with a $30 Amazon gift card and the fabricated results that were most difficult to detect received a bonus $50 Amazon gift card.

### Data analysis

```{r analysis study 1, echo=FALSE}
# Read in datafile, just in case the computations were not re-run.
dat_summary <- read.csv(dat_summary_file)
# Select only those from qualtrics who did for bonus
# removed <- sum(dat_summary$type == 'qualtrics' & is.na(dat_summary$bonus))
# dat_summary <- dat_summary[!(dat_summary$type == 'qualtrics' & is.na(dat_summary$bonus)), ]

type_index <- which(names(dat_summary) == 'type')
temp <- dat_summary[, c(type_index, 18:length(names(dat_summary)))]  
dat_summary$es_gender <- sqrt(dat_summary$es2_gender)
dat_summary$es_condition <- sqrt(dat_summary$es2_condition)
dat_summary$es_interaction <- sqrt(dat_summary$es2_interaction)
